---
title: "RRspread"
author: "Hsueh-Chien Raymond Cheng"
date: "05/02/2024"
output: html_document
---

```{r setup, include=FALSE, warning = FALSE, message = FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**This is the markdown file for generate results for Israel relative risk of spread manuscript**

### Load libraries

```{r, include=FALSE}
library(ape)
library(BactDating)
library(dplyr)
library(geosphere)
library(ggplot2)
library(data.table)
library(cowplot)
library(rrspread)

RR_matrix_notree = function(meta_table, coltime_range = 1,
                     des_distance_min = 0, des_distance_max = 50, ref_distance_min = 200, ref_distance_max = 300){
  lane.names <- meta_table$lane_id

  vector_lineage = meta_table$GPSC_PoPUNK2
  vector_location = meta_table$REGION
  vector_longitude <- meta_table$Longitude
  vector_latitude <- meta_table$Latitude
  vector_coltime = meta_table$Year_collection
  vector_coltime = vector_coltime + runif(length(vector_coltime), min = -0.01, max = 0.01)

  des_lineage_mat = abs(outer(vector_lineage, vector_lineage, "=="))
  diag(des_lineage_mat) <- NA
  colnames(des_lineage_mat) = lane.names
  rownames(des_lineage_mat) = lane.names
  ref_lineage_mat = 1- des_lineage_mat

  des_location_mat = abs(outer(vector_location, vector_location, "=="))
  diag(des_location_mat) <- NA
  colnames(des_location_mat) = lane.names
  rownames(des_location_mat) = lane.names
  ref_location_mat = 1 - des_location_mat

  coltime_mat = abs(outer(vector_coltime, vector_coltime, "-"))
  diag(coltime_mat) <- NA
  colnames(coltime_mat) = lane.names
  rownames(coltime_mat) = lane.names
  coltime_mat = (coltime_mat > 0)*(coltime_mat <= coltime_range) ## within a year

  des_distance_mat = gen_dist_mat(metadata = meta_table, min_dist = des_distance_min, max_dist = des_distance_max)
  ref_distance_mat = gen_dist_mat(metadata = meta_table, min_dist = ref_distance_min, max_dist = ref_distance_max)

  res_list = list("des_lineage_mat" = des_lineage_mat,
                  "ref_lineage_mat" = ref_lineage_mat,
                  "des_location_mat" = des_location_mat,
                  "ref_location_mat" = ref_location_mat,
                  "des_distance_mat" = des_distance_mat$distance_RRmatrix,
                  "ref_distance_mat" = ref_distance_mat$distance_RRmatrix,
                  "distance_num_mat" = ref_distance_mat$pairwise_distance_matrix,
                  "coltime_mat" = coltime_mat)
  return(res_list)
}

```


### Load input data
```{r}
metadata_file_path = "/Users/hc14/Documents/SpneumoIsrael/Israel_raw_data/input/Israel_1174isolates_metadata_input.txt"
metadata = read.csv(metadata_file_path, sep="\t")
metadata = metadata[, c("lane_id", "GPSC_PoPUNK2", "REGION", "Longitude", "Latitude", "Year_collection")]

```

### Load Gubbins tree

The tree should be recombination-corrected by Gubbins GPSC6,8,10,47,55 are selected for Israel migration analysis

```{r}
gpsc8_gubbin_tree_path = "/Users/hc14/Documents/SpneumoIsrael/analysis_results/05_topGPSC_gubbin_output_tree/GPSC8_reference_bwa.final_tree.tre"
  gpsc8_gubbin_tree = read.tree(gpsc8_gubbin_tree_path)

gpsc55_gubbin_tree_path = "/Users/hc14/Documents/SpneumoIsrael/analysis_results/05_topGPSC_gubbin_output_tree/GPSC55_reference_bwa.final_tree.tre"
  gpsc55_gubbin_tree = read.tree(gpsc55_gubbin_tree_path)

gpsc47_gubbin_tree_path = "/Users/hc14/Documents/SpneumoIsrael/analysis_results/05_topGPSC_gubbin_output_tree/GPSC47_reference_bwa.final_tree.tre"
  gpsc47_gubbin_tree = read.tree(gpsc47_gubbin_tree_path)

gpsc6_gubbin_tree_path = "/Users/hc14/Documents/SpneumoIsrael/analysis_results/05_topGPSC_gubbin_output_tree/GPSC6_reference_bwa.final_tree.tre"
  gpsc6_gubbin_tree = read.tree(gpsc6_gubbin_tree_path)

gpsc10_gubbin_tree_path = "/Users/hc14/Documents/SpneumoIsrael/analysis_results/05_topGPSC_gubbin_output_tree/GPSC10_reference_bwa.final_tree.tre"
  gpsc10_gubbin_tree = read.tree(gpsc10_gubbin_tree_path)
```

### Run bactdating (time-consuming)

**Input:** 1. Gubbins (recombination-corrected tree), 2. metadata with collection time   
**Output:** 1. bactdating results 2. root-to-tip distance vs. collection time   
**notes:**   
*drop_tip= TRUE* to make sure reference genome is not included in the analysis, *branch_model* can be assigned for evolutionary rates of branches, *n_it* to assign number of iterations in the MCMC process

```{r, include=FALSE}
gpsc6_gubbin_tree_plot = plt_gubbin_tree(gpsc6_gubbin_tree)
gpsc6_time_tree_list = match_tree_meta(metadata, gpsc6_gubbin_tree, "GPSC6_reference", drop_tip = TRUE)
gpsc6_bd_res = time_resolved(gpsc6_time_tree_list$match_tree_meta, gpsc6_time_tree_list$tree, "/Users/hc14/Documents/SpneumoIsrael/analysis_results/bd_res_GPSC6", n_it = 3000, branch_model="mixedgamma")



### or Load previous bactdating results
load(file = "/Users/hc14/Documents/SpneumoIsrael/analysis_results/bdtree_mixgamma.RData")
```

#### Checked convergence of MCMC

**notes:** 3 parameters (mu, sigma, alpha) should get converged and have enough sample size throughout MCMC. Each parameter should have sample size \> 100


```{r}
mcmc_gpsc6=as.mcmc.resBactDating(gpsc6_bd_res$BactDating) #check whether converge
coda::effectiveSize(mcmc_gpsc6)

```

### Bind trees and combine trees with metadata

match the laneid in the tips of the trees and the laneid in the metadata

```{r}
gpsc08_bd_tree = gpsc8_bd_res$BactDating$tree
gpsc55_bd_tree = gpsc55_bd_res$BactDating$tree
gpsc47_bd_tree = gpsc47_bd_res$BactDating$tree
gpsc06_bd_tree = gpsc6_bd_res$BactDating$tree
gpsc10_bd_tree = gpsc10_bd_res$BactDating$tree


multitrees = bind_multitree(gpsc08_bd_tree, gpsc55_bd_tree, gpsc47_bd_tree, gpsc06_bd_tree, gpsc10_bd_tree)
plot(multitrees, cex=0.3, no.margin = TRUE)

lanes <- multitrees$tip.label

## subset only those present in the trees
####metadata = read.csv(metadata_file)
selected_gpsc_metadata = subset(metadata, metadata$lane_id %in% lanes)
lanes_gpsc <- subset(lanes, lanes  %in% selected_gpsc_metadata$lane_id)
multitrees_overall <- keep.tip(multitrees, lanes_gpsc)

#Reorder tree based on order of tip labels
order <- multitrees_overall$tip.label
selected_gpsc_metadata <- selected_gpsc_metadata %>%
  dplyr::slice(match(order, lane_id))

```

### Load Metadata of GPSC6,8,10,47,55

```{r}
gpsc6_metadata <- selected_gpsc_metadata[which(selected_gpsc_metadata$GPSC_PoPUNK2 == "6"),]
gpsc8_metadata <- selected_gpsc_metadata[which(selected_gpsc_metadata$GPSC_PoPUNK2 == "8"),]
gpsc10_metadata <- selected_gpsc_metadata[which(selected_gpsc_metadata$GPSC_PoPUNK2 == "10"),]
gpsc47_metadata <- selected_gpsc_metadata[which(selected_gpsc_metadata$GPSC_PoPUNK2 == "47"),]
gpsc55_metadata <- selected_gpsc_metadata[which(selected_gpsc_metadata$GPSC_PoPUNK2 == "55"),]
```

### Generate matrices for relative risk calculations for each GPSC

**input:** 1. Metadata, 2. Bactdating results  
**output:** 11 matrices for following relative risk calculations

```{r}
gpsc6_rrmatrix <- RR_matrix(meta_table = gpsc6_metadata, bactdate_tree = gpsc6_bd_res$BactDating$tree, ref_distance_min = 0)
gpsc8_rrmatrix <- RR_matrix(meta_table = gpsc8_metadata, bactdate_tree = gpsc8_bd_res$BactDating$tree, ref_distance_min = 0)
gpsc10_rrmatrix <- RR_matrix(meta_table = gpsc10_metadata, bactdate_tree = gpsc10_bd_res$BactDating$tree, ref_distance_min = 0)
gpsc47_rrmatrix <- RR_matrix(meta_table = gpsc47_metadata, bactdate_tree = gpsc47_bd_res$BactDating$tree, ref_distance_min = 0)
gpsc55_rrmatrix <- RR_matrix(meta_table = gpsc55_metadata, bactdate_tree = gpsc55_bd_res$BactDating$tree, ref_distance_min = 0)
```

### Distance vs. Divergence time relationship

```{r}
### GPSC6
gpsc6_distance_dist <- as.data.frame(to.upper(gpsc6_rrmatrix$distance_num_mat))
gpsc6_divtime_dist <- as.data.frame(to.upper(gpsc6_rrmatrix$divtime_num_mat))
gpsc6_divtime_distance <- cbind(gpsc6_divtime_dist, gpsc6_distance_dist)
colnames(gpsc6_divtime_distance) <- c("divtime", "distance")

### GPSC8
gpsc8_distance_dist <- as.data.frame(to.upper(gpsc8_rrmatrix$distance_num_mat))
gpsc8_divtime_dist <- as.data.frame(to.upper(gpsc8_rrmatrix$divtime_num_mat))
gpsc8_divtime_distance <- cbind(gpsc8_divtime_dist, gpsc8_distance_dist)
colnames(gpsc8_divtime_distance) <- c("divtime", "distance")

### GPSC10
gpsc10_distance_dist <- as.data.frame(to.upper(gpsc10_rrmatrix$distance_num_mat))
gpsc10_divtime_dist <- as.data.frame(to.upper(gpsc10_rrmatrix$divtime_num_mat))
gpsc10_divtime_distance <- cbind(gpsc10_divtime_dist, gpsc10_distance_dist)
colnames(gpsc10_divtime_distance) <- c("divtime", "distance")

### GPSC47
gpsc47_distance_dist <- as.data.frame(to.upper(gpsc47_rrmatrix$distance_num_mat))
gpsc47_divtime_dist <- as.data.frame(to.upper(gpsc47_rrmatrix$divtime_num_mat))
gpsc47_divtime_distance <- cbind(gpsc47_divtime_dist, gpsc47_distance_dist)
colnames(gpsc47_divtime_distance) <- c("divtime", "distance")

### GPSC55
gpsc55_distance_dist <- as.data.frame(to.upper(gpsc55_rrmatrix$distance_num_mat))
gpsc55_divtime_dist <- as.data.frame(to.upper(gpsc55_rrmatrix$divtime_num_mat))
gpsc55_divtime_distance <- cbind(gpsc55_divtime_dist, gpsc55_distance_dist)
colnames(gpsc55_divtime_distance) <- c("divtime", "distance")

### GPSC all
gpsc_all_divtime_distance <- rbind(gpsc6_divtime_distance, gpsc8_divtime_distance, gpsc10_divtime_distance,gpsc47_divtime_distance, gpsc55_divtime_distance)
plot(gpsc_all_divtime_distance$divtime, gpsc_all_divtime_distance$distance)
```

### Distance vs. Divergence time with confidence interval by sampling the posterior bactdating trees

#### procedures:
Take each GPSC
1. Sample the posterior bactdating trees (we want to sample posterior trees sequentially through the MCMC trace)
2. Reconstruct the sampled posterior trees
3. Calculate mean pairwise evolutionary divergence time from all the sampled trees
4. Identify pairs within evolutionary divergence time
5. Sample pairs with replacement (bootstrap)
6. Extract spatial distance per evolutionary window
7. Calculate mean spatial  per evolutionary window

**Inputs:** 1. metadata, 2. bactdating results
```{r}
Dtime_Dist_CI_posterior_tree <- function(metadata, bd_res, initial_window=5, Ntreestosample=100, divtime_nboot=20, upper_window=50)
{
  ################################################################
  ## Sample trees from BactDating posterior
  ## Noemie Lefrancq
  ## Last update 21/09/2023
  ################################################################
  
  ## Load Bactdating result
  
  #resbd <- readRDS('bd_GPSC8')
  Ntips = length(bd_res$BactDating$tree$tip.label)
  Ntreestosample = 100
  ## ID of tree to sample
  tree_ID = round(seq(1, nrow(bd_res$BactDating$record), length.out = Ntreestosample))
  ################################################################
  ## Reconstruct one tree from the trace
  ################################################################
  ## Withdraw node heights
  
  posterior_sampling <- array(NA, dim=c(Ntips, Ntips, Ntreestosample))
  for(samp in 1:Ntreestosample){
    tip_dates = bd_res$BactDating$record[tree_ID[samp],1:Ntips]
    node_dates = bd_res$BactDating$record[tree_ID[samp],(Ntips+1):(2*Ntips-1)]
    tipandnodes_dates = c(tip_dates, node_dates)
    
    ## Take BD final tree for template
    reconstructed_tree = bd_res$BactDating$tree
    ## Replace edge lengths, based on the tip and node dates that were extracted above
    reconstructed_tree$edge.length = tipandnodes_dates[reconstructed_tree$edge[,2]] - tipandnodes_dates[reconstructed_tree$edge[,1]]
    ## pairwise distance
    reconstructed_pairwise_dist = cophenetic.phylo(reconstructed_tree)
    posterior_sampling[,,samp] <- reconstructed_pairwise_dist
  }
  
  for (i in 1:Ntips) {
    gpsc_divtime_posterior_mean <- apply(posterior_sampling, c(1,2), FUN=function(isamp) mean(isamp))
    gpsc_divtime_posterior_upperci <- apply(posterior_sampling, c(1,2), FUN = function(isamp) quantile(isamp, probs = 0.975, na.rm = T))
    gpsc_divtime_posterior_lowerci <- apply(posterior_sampling, c(1,2), FUN = function(isamp) quantile(isamp, probs = 0.075, na.rm = T))
  }
  
  gpsc_rrmatrix <- RR_matrix(meta_table = metadata, bactdate_tree = bd_res$BactDating$tree)
  gpsc_distance_dist <- as.data.frame(to.upper(gpsc_rrmatrix$distance_num_mat))
  gpsc_divtime_dist <- as.data.frame(to.upper(gpsc_divtime_posterior_mean))
  gpsc_dist_divtime_boot <- cbind(gpsc_distance_dist, gpsc_divtime_dist)
  colnames(gpsc_dist_divtime_boot) <- c("distance", "divtime")
  
  nboot = divtime_nboot
  x_divtime_max = seq(from=initial_window, to=upper_window, by=1)
  x_divtime_min = rep(0, length(x_divtime_max))
  #x_divtime_min = seq(from=0, to=20, by=1)
  
  boot_select <- data.frame()
  for (win in 1:length(x_divtime_max)) {
    boot.mean <- c()
    for (n in 1:nboot) {
      a = which(gpsc_dist_divtime_boot$divtime > x_divtime_min[win] &
                  gpsc_dist_divtime_boot$divtime < x_divtime_max[win])        #select a certain region (index)
      if (length(a)!=0){
        b = sample(x=a, size=100, replace = T)  #subsample with replacement
        resample.mean <- mean(gpsc_dist_divtime_boot[b, ]$distance)
        boot.mean <- c(boot.mean, resample.mean)
      }
    }
    distance.med <- quantile(boot.mean, probs = c(0.5), na.rm=T) #calculate confidence interval
    distance.mean <- mean(boot.mean)
    distance.upperci <- quantile(boot.mean, probs = c(0.975), na.rm=T)
    distance.lowerci <- quantile(boot.mean, probs = c(0.025), na.rm=T)
    mid_win <- mean(c(x_divtime_min[win], x_divtime_max[win]))
    window_divtime_distance <- cbind(x_divtime_min[win], x_divtime_max[win], mid_win, distance.med, distance.mean, distance.upperci, distance.lowerci)
    boot_select <- rbind(boot_select, window_divtime_distance)
  }
  
  res_bootstrap_divtime_dist = list("window_divtime_dist" = boot_select, "divtime_dist_table" = gpsc_dist_divtime_boot)
  return(res_bootstrap_divtime_dist)
}
```

#### run Dtime_Dist_CI_posterior_tree for each GPSC
````{r}
gpsc6_Dtime_Dist_CI_posterior_tree <- Dtime_Dist_CI_posterior_tree(gpsc6_metadata, gpsc6_bd_res)
#gpsc8_Dtime_Dist_CI_posterior_tree <- Dtime_Dist_CI_posterior_tree(gpsc8_metadata, gpsc8_bd_res)
#gpsc10_Dtime_Dist_CI_posterior_tree <- Dtime_Dist_CI_posterior_tree(gpsc10_metadata, gpsc10_bd_res)
#gpsc47_Dtime_Dist_CI_posterior_tree <- Dtime_Dist_CI_posterior_tree(gpsc47_metadata, gpsc47_bd_res, initial_window=10)
#gpsc55_Dtime_Dist_CI_posterior_tree <- Dtime_Dist_CI_posterior_tree(gpsc55_metadata, gpsc55_bd_res)
````

#### plot Distance CI vs. Divergence (GPSC6)
```{r}
gpsc6_dist_div <- ggplot(gpsc6_Dtime_Dist_CI_posterior_tree$window_divtime_dist, aes(x=mid_win, y=distance.mean, ymin=distance.lowerci, ymax=distance.upperci))+
  geom_line()+
  geom_ribbon(fill="pink", alpha=0.5)+
  #xlim(0,60)+
  xlab("Divergence Time \n (years)")+
  ylab("Distance (km)")+
  ylim(c(0,87))+
  theme(panel.background = element_blank(), axis.line = element_line(colour = "black"), text = element_text(family = "Helvetica"))

gpsc6_dist_div

```

### Relative risk of Spread: lineage v.s. district
```{r message=FALSE, include=FALSE, warning=FALSE, paged.print=FALSE}
rr_lineage_location_mat = RR_matrix(selected_gpsc_metadata, multitrees)

colyear_mat_ones = matrix(1, nrow=dim(selected_gpsc_metadata)[1], ncol=dim(selected_gpsc_metadata)[1])  ## don't select by collection year
rr = RR_bootstrap(metadataset = selected_gpsc_metadata,
                  des_geo_matrix = rr_lineage_location_mat$des_location_mat,
                  ref_geo_matrix = rr_lineage_location_mat$ref_location_mat,
                  des_strain_matrix = rr_lineage_location_mat$des_lineage_mat,
                  ref_strain_matrix = rr_lineage_location_mat$ref_lineage_mat,
                  colyear_mat_ones, nboot = 100, subsample_min = 70)
## load library: ggplot2
district_lineage_rr = PltRR(rr, "Within District Relative Ratio", "District")

# metadata: metadata
lineage_matrix <- RR_matrix_notree(meta_table = metadata)

colyear_mat_ones = matrix(1, nrow=dim(metadata)[1], ncol=dim(metadata)[1])  ## don't select by collection year
rr = RR_bootstrap(metadataset = metadata,
                  des_geo_matrix = lineage_matrix$des_location_mat,
                  ref_geo_matrix = lineage_matrix$ref_location_mat,
                  des_strain_matrix = lineage_matrix$des_lineage_mat,
                  ref_strain_matrix = lineage_matrix$ref_lineage_mat,
                  colyear_mat_ones, nboot = 100, subsample_min = 70)

## load library: ggplot2
district_lineage_rr = PltRR(rr, "Within District Relative Ratio", "District")
district_lineage_rr
```
```{r}
district_lineage_rr
```


### Relative risk of Spread: lineage v.s. spatial distance
rolling window from 0-30 to 120-150 km 

```{r message=FALSE, include=FALSE, warning=FALSE, paged.print=FALSE}
des_distance_min1 = c(0,1)
des_distance_min2 = seq(from=5, to=25, by=5)
des_distance_min3 = seq(from=30, to=90, by=10)
des_distance_min = append(append(des_distance_min1, des_distance_min2), des_distance_min3)
des_distance_max1 = c(1,5)
des_distance_max2 = seq(from=10, to=30, by=5)
des_distance_max3 = seq(from=40, to=100, by=10)
des_distance_max = append(append(des_distance_max1, des_distance_max2), des_distance_max3)

des_distance_range = length(des_distance_max)

res_list = c()
res_df = data.frame()

colyear_mat_ones = matrix(1, nrow=dim(metadata)[1], ncol=dim(metadata)[1])
diag(colyear_mat_ones) <- NA

RRplotLists = list()
RRdistanceLists = list()
RRdframLists = list()

for(distances in 1:des_distance_range){
  distance_name = paste0("Distance: ",des_distance_min[distances],"~",des_distance_max[distances],"km")
  dist_lineage_mat = RR_matrix_notree(meta_table = metadata, 
                               des_distance_min = des_distance_min[distances],
                               des_distance_max = des_distance_max[distances], 
                               ref_distance_min = 50, ref_distance_max = 300)
  dist_lineage_mat_rr = RR_bootstrap(metadataset = metadata,
                                     des_geo_matrix = dist_lineage_mat$des_distance_mat,
                                     ref_geo_matrix = dist_lineage_mat$ref_distance_mat,
                                     des_strain_matrix = dist_lineage_mat$des_lineage_mat,
                                     ref_strain_matrix = dist_lineage_mat$ref_lineage_mat,
                                     colyear_matrix = colyear_mat_ones,
                                     nboot = 100,
                                     subsample_min = 100)
  rr_plot = PltRR(dist_lineage_mat_rr, distance_name, distance_name)
  RRplotLists[[distances]] = print(rr_plot)
  
  distances_col = data.frame(distance_min=des_distance_min[distances], distance_max=des_distance_max[distances], distance_mid = (des_distance_min[distances]+des_distance_max[distances])/2)
  new_row = cbind(dist_lineage_mat_rr, distances_col)
  res_df = rbind(res_df, new_row)
  
  assign(distance_name, rr_plot)
  res_list=c(res_list, distance_name)
}

```


### Figure 2: Combine plots 1. lineage vs. district & plot 2. lineage vs. spatial distance
```{r}
dist_district_lineage_rr_table <- rbind(res_df, rr, fill=TRUE)
for(i in 1:nrow(dist_district_lineage_rr_table)){
  dist_district_lineage_rr_table[i, "distance_window"] = paste0(dist_district_lineage_rr_table[i, "distance_min"],"-" ,dist_district_lineage_rr_table[i, "distance_max"])
}
dist_district_lineage_rr_table[nrow(dist_district_lineage_rr_table), "distance_window"] = "within \ndistrict"
dist_district_lineage_rr_table$distance_window = as.factor(dist_district_lineage_rr_table$distance_window)


#png(filename = "/Users/hc14/Documents/SpneumoIsrael/analysis_results/figures/rolling_distance_RR.png", res=1200, width = 5000, height = 3000)
dist_district_lineage_rr_table$distance_window <- ordered(dist_district_lineage_rr_table$distance_window, levels=c("0-1", "1-5", "5-10", "10-15", "15-20", "20-25", "25-30", "30-40", "40-50", "50-60", "60-70", "70-80", "80-90", "90-100", "within \ndistrict"))
rolling_distance_RR_plot_50km <- ggplot(dist_district_lineage_rr_table, aes(x=distance_window, y=RR_mean))+
  geom_pointrange(data=dist_district_lineage_rr_table, mapping=aes(x=as.factor(distance_window), y=RR_mean, ymin = RR_lower_ci, ymax = RR_upper_ci), alpha=0.9) +
  geom_hline(yintercept=1, linetype="dashed", color = "red") + ##dashed line at 1
  #scale_y_log10()+
  ylab("Risk Ratio")+
  xlab("Distance window (km)")+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), axis.line = element_line(colour = "black"))+
  geom_rect(xmin = 14.5, xmax = 15.5, ymin = -1,  ymax = 5,
            fill = 'blue', alpha = 0.01) +
  geom_rect(xmin = 0, xmax = 15.5, ymin = -1, ymax = 5,
            fill = 'red', alpha = 0.01)

rolling_distance_RR_plot_50km

```


### Rolling divergence vs. relative ratio | within district vs. between district

```{r, include=FALSE}
des_divtime_min1 = seq(from=0, to=8, by=2)
des_divtime_min2 = seq(from=10, to=15, by=5)
des_divtime_min = c(des_divtime_min1, des_divtime_min2, 20)
des_divtime_max1 = seq(from=2, to=10, by=2)
des_divtime_max2 = seq(from=15, to=20, by=5)
des_divtime_max = c(des_divtime_max1, des_divtime_max2, 30)
des_divtime_range = length(des_divtime_max)

res_list = c()
res_df_divtime = data.frame()

colyear_mat_ones = matrix(1, nrow=dim(selected_gpsc_metadata)[1], ncol=dim(selected_gpsc_metadata)[1])
diag(colyear_mat_ones) <- NA

RRplotLists = list()
RRdivtimeLists = list()
RRdframLists = list()

for(divtimes in 1:des_divtime_range){
  divtime_name = paste0("Divergence: ",des_divtime_min[divtimes],"-",des_divtime_max[divtimes],"yrs")
  loc_divtime_mat = RR_matrix(meta_table = selected_gpsc_metadata, bactdate_tree = multitrees,
                              des_divtime_min = des_divtime_min[divtimes],
                              des_divtime_max = des_divtime_max[divtimes])
  loc_divtime_mat_rr = RR_bootstrap(metadataset = selected_gpsc_metadata,
                                    des_geo_matrix = loc_divtime_mat$des_location_mat,
                                    ref_geo_matrix = loc_divtime_mat$ref_location_mat,
                                    des_strain_matrix = loc_divtime_mat$des_divtime_mat,
                                    ref_strain_matrix = loc_divtime_mat$ref_divtime_mat,
                                    colyear_matrix = colyear_mat_ones,
                                    nboot = 100,
                                    subsample_min = 70)
  rr_plot = PltRR(loc_divtime_mat_rr, divtime_name, divtime_name)
  RRplotLists[[divtimes]] = print(rr_plot)
  
  divtimes_col = data.frame(divtime_min=des_divtime_min[divtimes], divtime_max=des_divtime_max[divtimes], divtime_mid = (des_divtime_min[divtimes]+des_divtime_max[divtimes])/2)
  new_row = cbind(loc_divtime_mat_rr, divtimes_col)
  res_df_divtime = rbind(res_df_divtime, new_row)
  
  assign(divtime_name, rr_plot)
  res_list=c(res_list, divtime_name)
}

#write.table(res_df_divtime, "rolling_distance_divtime_district.txt", sep = "\t", quote = FALSE, row.names =FALSE, col.names = TRUE)

rolling_distance_divtime <- PltRR_rolling_divtime(res_df_divtime,"rolling divergence")+
  scale_x_discrete(labels = c("0-2", "2-4", "4-6", "6-8", "8-10", "10-15", "15-20", "20-30"))+
  #scale_y_continuous(breaks = c(0.5, 1, 1.5, 2, 2.5, 3, 3.5, 4, 4.5, 5), labels = c("0.5", "1.0", "1.5", "2.0", "2.5", "3.0", "3.5", "4.0", "4.5", "5.0"))+
  scale_y_log10(limits=c(0.1,7))+
  ggtitle("By District")+
  theme(axis.text = element_text(size=18),title = element_text(size=20))
```

### Rolling divergence vs. relative ratio | close area (0-30 km) vs. distant area

```{r, include=FALSE}
des_divtime_min1 = seq(from=0, to=8, by=2)
des_divtime_min2 = seq(from=10, to=15, by=5)
des_divtime_min = c(des_divtime_min1, des_divtime_min2, 20)
des_divtime_max1 = seq(from=2, to=10, by=2)
des_divtime_max2 = seq(from=15, to=20, by=5)
des_divtime_max = c(des_divtime_max1, des_divtime_max2, 30)
des_divtime_range = length(des_divtime_max)

res_list = c()
res_df = data.frame()

colyear_mat_ones = matrix(1, nrow=dim(selected_gpsc_metadata)[1], ncol=dim(selected_gpsc_metadata)[1])
diag(colyear_mat_ones) <- NA

RRplotLists = list()
RRdivtimeLists = list()
RRdframLists = list()

for(divtimes in 1:des_divtime_range){
  divtime_name = paste0("Divergence: ",des_divtime_min[divtimes],"~",des_divtime_max[divtimes],"yrs")
  loc_divtime_mat = RR_matrix(meta_table = selected_gpsc_metadata, bactdate_tree = multitrees,
                              des_divtime_min = des_divtime_min[divtimes],
                              des_divtime_max = des_divtime_max[divtimes],
                              des_distance_min = 0,
                              des_distance_max = 30, 
                              ref_distance_min = 150,
                              ref_distance_max = 300)
  loc_divtime_mat_rr = RR_bootstrap(metadataset = selected_gpsc_metadata,
                                    des_geo_matrix = loc_divtime_mat$des_distance_mat,
                                    ref_geo_matrix = loc_divtime_mat$ref_distance_mat,
                                    des_strain_matrix = loc_divtime_mat$des_divtime_mat,
                                    ref_strain_matrix = loc_divtime_mat$ref_divtime_mat,
                                    colyear_matrix = colyear_mat_ones,
                                    nboot = 100,
                                    subsample_min = 70)
  rr_plot = PltRR(loc_divtime_mat_rr, divtime_name, divtime_name)
  RRplotLists[[divtimes]] = print(rr_plot)
  
  divtimes_col = data.frame(divtime_min=des_divtime_min[divtimes], divtime_max=des_divtime_max[divtimes], divtime_mid = (des_divtime_min[divtimes]+des_divtime_max[divtimes])/2)
  new_row = cbind(loc_divtime_mat_rr, divtimes_col)
  res_df = rbind(res_df, new_row)
  
  assign(divtime_name, rr_plot)
  res_list=c(res_list, divtime_name)
}

#write.table(res_df, "/Users/hc14/Documents/SpneumoIsrael/analysis_results/figures/figure4/rolling_distance_divtime_30km_150kmRef.txt", sep = "\t", quote = FALSE, row.names =FALSE, col.names = TRUE)

rolling_distance_divtime_30km_150kmRef <- PltRR_rolling_divtime(res_df,"rolling divergence")+
  scale_x_discrete(labels = c("0-2", "2-4", "4-6", "6-8", "8-10", "10-15", "15-20", "20-30"))+
  #scale_y_continuous(breaks = c(0.5, 1, 1.5, 2, 3, 4, 5, 6, 7), labels = c("0.5", "1.0", "1.5", "2.0", "3.0",  "4.0", "5.0", "6.0", "7.0"))+
  scale_y_log10(limits=c(0.1,60))+
  ggtitle("By Distance")+
  theme(axis.text = element_text(size=18),title = element_text(size=20))

#png("/Users/hc14/Documents/SpneumoIsrael/analysis_results/figures/figure4/rolling_divtime_30km_150kmRef.png")
#rolling_distance_divtime_30km_150kmRef
#dev.off()
```

### Figure 4: rolling divergence vs. relative risk ratio

```{r}
rolling_divergence_rr <- plot_grid(rolling_distance_divtime, NULL,rolling_distance_divtime_30km_150kmRef, rel_widths=c(1,0,1),rel_heights=c(1,0,1), scale=c(1,0,1), align = "hv", nrow = 1)

#png("/Users/hc14/Documents/SpneumoIsrael/analysis_results/figures/figure4/figure4_30km_100kmRef.png", width = 1000, height = 300)
rolling_divergence_rr
#dev.off()


#rolling_divergence_rr_sensitive <- plot_grid(rolling_distance_divtime_30km_0kmRef,rolling_distance_divtime_30km_50kmRef, rolling_distance_divtime_30km_80kmRef, rel_widths=c(1,1,1),rel_heights=c(1,1,1), scale=c(1,1,1), align = "hv", nrow = 2, labels = "AUTO")

#png("/Users/hc14/Documents/SpneumoIsrael/analysis_results/figures/figure4/rolling_divergence_rr_sensitive.png", width = 1600, height = 800)
#rolling_divergence_rr_sensitive
#dev.off()
```





